---
title: "Merging patient entries in COVID data"
output: html_notebook
---
```{r}
library(tidyverse)
library(ggplot2)
library(ggsci) 
library(lubridate)
library(corrplot)
library(data.table)
library(rstan)
library(cowplot)
library(scales)
```

```{r}
processed_data_2 <- read_csv("processed_with_mutations_na_as_no_vaccination_processed_073021.csv")
```
First we want to kill off individuals who don't even have a CPR.

```{r}
cat(sprintf("%i patient entries missing CPR", sum(is.na(processed_data_2$CPR))))
processed_data_2 <- dplyr::filter(processed_data_2, !is.na(CPR))
processed_data_2 <- processed_data_2[,-(1:5)]
```
We also have genotypes of individuals who weren't sequenced coded as N* and we should set these to NA.
```{r}
rbind(56:60, colnames(processed_data_2[,56:60]))
```
```{r}
rbind(1410:1413, colnames(processed_data_2[,1410:1413]))
```
So we want to NA everything from 57 to 1211. This is because we are assuming that Justin added genotype information to entries only when a lab number was present.
```{r}
processed_data_2[is.na(processed_data_2$`Lab No`), 57:1411] <- NA
```
Write a simple function to merge rows and complain when that merging contains contradictions.
```{r}
merge.rows <- function(rowlist){
  result <- processed_data_2[1,]
  for(ii in 1:ncol(rowlist))
    result[1,ii] <- NA
  for(ii in 1:ncol(rowlist)){
    if((names(rowlist[,ii])=="Patient Id") | (names(rowlist[,ii])=="CPR")){
      CPR.nums <- unique(as.numeric(unlist(rowlist[,ii])))
      CPR.nums <- CPR.nums[!is.na(CPR.nums)]
      if(length(CPR.nums)>1)
        print(paste0("multiple CPR: ", CPR.nums))
      result[ii] <- CPR.nums[1]
    } else{
      vals <- unlist(rowlist[,ii])
      nonmiss.vals <- unique(vals[!is.na(vals)])
      if(length(nonmiss.vals)==0)
        next
      if(length(nonmiss.vals)>1){
        print("multiple vals:")
        print(c(names(rowlist[,ii]), nonmiss.vals))
        if(names(rowlist[,ii])=="Age"){
          result[ii] <- min(nonmiss.vals)
        } else if((names(rowlist[,ii])=="Vaccination Status") | (names(rowlist[,ii])=="Vaccination Status (Processed)")){
          if("Yes" %in% nonmiss.vals)
            result[ii] <- "Yes"
          else
            result[ii] <- "No"
        }
      }
      if(grepl("Date", names(rowlist)[ii])){
        result[ii] <- lubridate::as_date(nonmiss.vals[1])
      } else{
         result[ii] <- nonmiss.vals[1]
      }
    }
  }
  return(result)
}
```
Run this with some test rows to see if it is working properly.
```{r}
rowlist <- processed_data_2[c(28, 126, 261),1:10]
merge.test <- merge.rows(rowlist)
```

```{r}
rowlist
merge.test
```

```{r}
matching.sets <- list()
jj <- 1
for(ii in 1:nrow(processed_data_2)){
  CPR <- processed_data_2$CPR[ii]
  labno <- processed_data_2$`Lab No`[ii]
  matching.rows <- grepl(CPR, processed_data_2$CPR) | grepl(CPR, processed_data_2$`Lab No`)
  if(sum(matching.rows)> 1){
    matching.sets[[jj]] <-  which(matching.rows)
    jj <- jj + 1
  }
}
```

```{r}
merged.matches <- list(-1)
for(matching.set in matching.sets){
  found.matches <- c()
  for(ii in 1:length(merged.matches)){
    if(length(base::intersect(matching.set, merged.matches[[ii]]))>0){
      found.matches <- c(found.matches, ii)
    }
  }
  if(length(found.matches)==0){
    merged.matches[[length(merged.matches)+1]] <- matching.set
  } else{
    new.set <- matching.set
    for(jj in found.matches){
      new.set <- c(new.set, merged.matches[[jj]])
    }
    new.set <- unique(new.set)
    merged.matches <- merged.matches[-found.matches]
    merged.matches[[length(found.matches)+1]] <- new.set
  }
}
merged.matches[[1]] <- NULL
```

```{r}
all.matched <- do.call(c, merged.matches)
length(all.matched) == length(unique(all.matched))
```

```{r}
merged.rows <- list()
ii <- 1
for(merged.match in merged.matches){
  print(c("merging", merged.match))
  merged.rows[[ii]] <- merge.rows(processed_data_2[merged.match,])
  ii <- ii + 1
}
```
```{r}
merge.test <- do.call(rbind, merged.rows)
```

Now we need to remove the dupcliate rows from the table and add the merged rows in.
```{r}
merged_data <- processed_data_2[-all.matched,] 
for(ii in 1:length(merged.rows)){
  merged_data <- rbind(merged_data, merged.rows[[ii]])
}
```
Now read in the nextclade data
```{r}
nextclade <- fread("nextclade.tsv", sep="\t", fill=T)
nextclade$seqName <- sapply(nextclade$seqName, function(X){
  if(grepl("_", X)){
    return(str_split(X, "_")[[1]][1])
  } else{
    return(X)
  }
})
```


```{r}
colnames(merged_data)[1:60]
```
We're going to add the nextclade data ourselves just to make sure the lab IDs are matched correctly to records now.
```{r}
merged_data$qc.overallStatus <- NA
merged_data$totalMissing <- NA
merged_data$totalSubstitutions <- NA
merged_data$totalPcrPrimerChanges <- NA
merged_data$qc.overallScore <- NA
merged_data$qc.missingData.status <- NA
merged_data$qc.mixedSites.status <- NA
merged_data$qc.snpClusters.status <- NA
merged_data$clade <- NA

for(ii in 1:nrow(nextclade)){
  if(ii == 695) ## Skip 695 because it is a lower-quality duplicate... see below for reasoning
    next
  seqname <- as.character(nextclade$seqName[ii])
  qc.overallStatus <- nextclade$qc.overallStatus[ii]
  qc.overallScore <- nextclade$qc.overallScore[ii]
  totalMissing <- nextclade$totalMissing[ii]
  totalSubstitutions <- nextclade$totalSubstitutions[ii]
  totalPcrPrimerChanges <- nextclade$totalPcrPrimerChanges[ii]
  qc.missingData.status <- nextclade$qc.missingData.status[ii]
  qc.mixedSites.status <- nextclade$qc.mixedSites.status[ii]
  qc.snpClusters.status <- nextclade$qc.snpClusters.status[ii]
  clade <- nextclade$clade[ii]
  matching.rows <- grepl(seqname, merged_data$`Lab No`) | grepl(seqname, merged_data$CPR)
  if(sum(matching.rows)>1){
    print(c(ii, seqname, merged_data$CPR[matching.rows]))
    print(which(matching.rows))
  }
  merged_data$qc.overallStatus[matching.rows] <- qc.overallStatus
  merged_data$qc.missingData.status[matching.rows] <- qc.missingData.status
  merged_data$qc.mixedSites.status[matching.rows] <- qc.mixedSites.status
  merged_data$qc.snpClusters.status[matching.rows] <- qc.snpClusters.status
  merged_data$qc.overallScore[matching.rows] <- qc.overallScore
  merged_data$totalMissing[matching.rows] <- totalMissing
  merged_data$totalSubstitutions[matching.rows] <- totalSubstitutions
  merged_data$totalPcrPrimerChanges[matching.rows] <- totalPcrPrimerChanges
  merged_data$clade[matching.rows] <- clade
}
```

```{r}
other_comorbidities <- vector(mode="character", length=nrow(seq_data))
is_yes <- function(value) {
  if (is.na(value)) {
    return(FALSE)
  } else if (value == "Yes") {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

for (i in 1:nrow(seq_data)) {
  if (
    is_yes(seq_data$`Immunocompromizing.Disease`[i]) |
    is_yes(seq_data$`Renal.Disease`[i]) |
    is_yes(seq_data$Obesity[i]) |
    is_yes(seq_data$`Chronic.Lung.Disease`[i]) |
    is_yes(seq_data$`Cardiovascular.Disease`[i]) |
    is_yes(seq_data$`Sickle.Cell.Disease.Or.Trait`[i]) |
    is_yes(seq_data$Cancer[i])
  ) {
    other_comorbidities[i] <- "Yes";
  } else {
    other_comorbidities[i] <- "No";
  }
}

seq_data["Other Comorbidities"] <- other_comorbidities
```

```{r}
seq_data$day <- lubridate::day(seq_data$`Collection.Date`) + 
    (lubridate::month(seq_data$`Collection.Date`)-1)*30.5 + 
    (lubridate::year(seq_data$`Collection.Date`)-2020)*365
seq_data$day <- seq_data$day - min(seq_data$day, na.rm = T)
seq_data$region.processed <- seq_data$Region
seq_data$region.processed[is.na(seq_data$Region)] <- "None"
```

```{r}
kill.row <- (seq_data$day < 200 | seq_data$day > 400) & seq_data$clade == "20I (Alpha, V1)"
kill.row[is.na(kill.row)] <- FALSE
seq_data <- seq_data[!kill.row,]
```

```{r}
clade.meds <- group_by(seq_data, clade) %>% summarize(med.day=median(day, na.rm=T))
rownames(clade.meds) <- clade.meds$clade
seq_data$day_imputed <- seq_data$day
for(ii in which(is.na(seq_data$day))){
  seq_data$day_imputed[ii] <- clade.meds[seq_data$clade[ii],]$med.day
}
```

```{r}
seq_data$age_imputed <- seq_data$Age
seq_data$age_imputed[is.na(seq_data$Age)] <- median(seq_data$Age, na.rm=T)
seq_data$Gender.num <- as.numeric(seq_data$Gender=="Female")
seq_data$Gender.num[is.na(seq_data$Gender)] <- mean(seq_data$Gender.num, na.rm=T) 
```


```{r}
seq_data$Nationality_processed <- "Other"
seq_data$Nationality_processed[seq_data$Nationality=="Bahraini"] <- "Bahraini"
seq_data$Nationality_processed[seq_data$Nationality %in% c("Indian", "Sri Lankan", "Nepalese", "Nepalise", "Bangladeshi")] <- "S_Asia"
seq_data$Nationality_processed[is.na(seq_data$Nationality)] <- "None"

seq_data_hq$Nationality_processed <- "Other"
seq_data_hq$Nationality_processed[seq_data_hq$Nationality=="Bahraini"] <- "Bahraini"
seq_data_hq$Nationality_processed[seq_data_hq$Nationality %in% c("Indian", "Sri Lankan", "Nepalese", "Nepalise", "Bangladeshi")] <- "S_Asia"
seq_data_hq$Nationality_processed[is.na(seq_data_hq$Nationality)] <- "None"
```

```{r}
seq_data$qc.missingData.status <- factor(seq_data$qc.missingData.status, c("bad", "mediocre", "good"))
```


```{r}
seq_data_hq <- dplyr::filter(seq_data, qc.missingData.status!="bad")
```

```{r}
write.table(seq_data, file="seq_data_082721.csv", row.names = F, sep=",")
write.table(seq_data_hq, file="seq_data_hq_082721.csv", row.names = F, sep=",")
```




